---
title: "Exploring SQM"
author: "Mark Heckmann"
date: "31 5 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Input

Using the grid from p. 14:

```{r echo = FALSE, message = FALSE, warning = FALSE}
library(OpenRepGrid)
library(tidyverse)

x <- importExcel("gallifa-botella-grid.xlsx")
bertin(x)
```


## STEP 2: INTEGRATION AND DIFFERENTIATION INDEXES

```{r echo=FALSE}

# understanding Step III:

# Similarity matrix for element "Freud"
# elements
# constructs                Adler Jung Rogers Piaget Vygotsky Skinner Pavlov
# Psychoanal - Non-psycho     4    4      1      0        0       0      0
# Psychoanal - Constructi     3    4      1      1        2       2      2
# Behavioral - Psychoanal     4    4      2      2        2       0      1
# Developmen - Behavioral     3    3      3      4        4       1      1
# Psychoanal - Personalit     4    4      1      1        2       2      2
# Behavioral - Psycholing     4    4      3      3        2       2      2
# Behavioral - Developmen     3    3      4      4        4       1      1

# create similarity matrix as grid
similarity_grid <- function(x, e = 1)
{
  k <- getScale(x) %>% diff %>% unname

  s <- k - abs(getRatingLayer(x) - getRatingLayer(x[, rep(e, ncol(x))]))
  s_e <- x[, -e]
  s_e[, ] <- 1
  s_e <- setScale(s_e, min = 0, max = 4)
  s_e[, ] <- s[, -e]
  
  leftpoles(s_e) <- paste(" —", leftpoles(x),  "/", rightpoles(x))
  rightpoles(s_e) <- paste(" +", rightpoles(x), "/", leftpoles(x))
  s_e
}
```

The similarity matrix from step I in the paper is the following for element 1 = Freud.

```{r echo=FALSE}
e <- 1
s_e <- similarity_grid(x, e = e)
bertin(s_e)
```

Visualizing the similarities as a biplot using midpooint centering. We can read off, for example, that the similarity of "Adler" and "Jung" to the base element "Freud" is high on the psychoanalytic / behavior construct (indicated by the "+"). Conversely, we see that the similarity between "Pavlov" / "Skinner" and "Freud" is low on this construct. **However, it remains unclear how construct loadings could indicate this in any way for all elements.**

```{r echo=FALSE, fig.width = 8}
biplot2d(s_e, mai = c(.2, 2.5, .2, 2.5), center = 4, c.point.cex = 0)
title(paste("Similarity to", elements(x)[e]))

# add construct vectors for testing purposes
r <- getRatingLayer(s_e)
d <- svd(r - getScaleMidpoint(s_e)) # 4 = midpoint centering
points(d$u[ , 1:2] * 3, pch=4) 

```

```{r echo = FALSE}
round(d$u %*% diag(d$d) %*% t(d$v))
```





